Extracting sunbeams from cucumbers: Computational reproducibility is essential for archaeological science

### 2025-02-03 paper.qmd

At the time of writing (Feburary 2025) I have completed 47 reproducibility reviews of 25 manuscripts submitted to JAS. Of these, 11 have been published in JAS to date. Seven of these eleven papers fully passed the reproducibility review, resulting in a success rate, by one measure, of 63%. Four of the seven papers could be fully reproduced on my first attempt, the others required additional input from the authors. For comparison with reproducibility studies in other fields reported above, the seven fully reproducible papers should be divided by the 25 reviewed for reproducibility, resulting in a 28% success rate. Expanding the denominator to include the total number of research articles published in JAS from May 2024 (when the first article to pass the reproducibility review, @herskindComputationalLinguisticMethodology2024, was published) to January 2025 (n = 97), which are articles that could have been eligible for reproducibility review had the authors used an open source programming language (e.g. instead of commercial software such as Microsoft Excel or SPSS, etc.) the success rate is 7%. In any case, the computational reproducibility of archaeological research is generally on the low end of the distribution of values obtained from a variety of harder and soft sciences.

>>>>

At the time of writing (January 2025) I have completed 47 reproducibility reviews of 25 manuscripts submitted to JAS. Of these, 11 have been published in JAS to date. Seven of these eleven papers fully passed the reproducibility review, resulting in a success rate, by one measure, of 63%. Four of the seven papers could be fully reproduced on my first attempt, the others required additional input from the authors. For comparison with reproducibility studies in other fields reported above, the seven fully reproducible papers should be divided by the 25 reviewed for reproducibility, resulting in a 28% success rate. Expanding the denominator to include the total number of research articles published in JAS from May 2024 (when the first article to pass the reproducibility review, @herskindComputationalLinguisticMethodology2024, was published) to January 2025 (n = 97), which are articles that could have been eligible for reproducibility review had the authors used an open source programming language (e.g. instead of commercial software such as Microsoft Excel or SPSS, etc.) the success rate is 7%. In any case, the computational reproducibility of archaeological research is generally on the low end of the distribution of values obtained from a variety of hard and soft sciences.
